Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-10 manuscript.tex
update fig3 resolution

\textbf{Macrostate clustering.} The 200 microstates comprising each system were finally coarsened into our terminal macrostate MSM. The microstate transition matrix for each system was computed at a range of lag times $\tau$ and then diagonalized to recover the corresponding eigenvalues $\lambda_i$ and associated implied time scales $t_i = -\tau / \ln \left| \lambda i \right|$.~\citep{Wehmeyer2019IntroductionV1.0} The implied time scale plots for the four DNA sequences are presented in SI Figure 2. We observe convergence of the implied time scales very rapidly with lag time $\tau$ for all systems, motivating the construction of high resolution macrostate MSMs at a lag time $\tau$ = 1.2 ns. At this choice of lag time, we recover 5, 4, 2, and 1 implied time scales for the AT-all, GC-end, GC-core, and GC-mid systems, respectively. The identification of $(i-1)$ implied time scales \noteb{(may be 1-2 additional converging, but these correspond to diffusive processes that are irrelevant or unhelpful in determining the hybridization mechanisms. Still, this might come up when looking at the the SI Fig 2} implies the presence of $(i-1)$ slow modes and motivates the coarsening of the system into $i$ macrostates. We estimate these $i$ macrostates by applying PCCA+ spectral clustering to the leading $(i-1)$ eigenvectors of the microstate transition matrix.~\citep{Roblitz2013FuzzyClassification, Weber2018ImplicationsSimulation, Kube2007AConformations} We then estimate the corresponding 6, 5, 3, and 2 macrostate transition matrices $\mathbf{P}$ for the AT-all, GC-end, GC-core, and GC-mid systems, respectively, by projecting the MD simulation trajectories into these discrete macrostates. These macrostate MSMs constitute our terminal kinetic models. We validate the Markovian nature of the four MSMs by subjecting them to the Chapman-Kolmogorov (CK) test \citep{Noe2009ConstructingSimulations,Phys2011MarkovValidation,  Wehmeyer2019IntroductionV1.0}. This test asserts that the transition matrix for a Markovian (i.e., memoryless) MSM constructed at a lag time $\tau$ should satisfy the condition $\mathbf{P}(k \tau) = \mathbf{P}^k(\tau)$, which states that $k$ successive applications of the transition matrix constructed at a lag time $\tau$should be equivalent to a single application of the transition matrix constructed at a lag time $k \tau$. We present in SI Figure 3 the CK tests for each DNA sequence to demonstrate that the $\tau$ = 1.2 ns models performs excellently in predicting transition probabilities out to $k \tau$ = 7.2 ns, validating the Markovian nature and kinetic validity of the four kinetic models.

>>>>

\textbf{Macrostate clustering.} The 200 microstates comprising each system were finally coarsened into our terminal macrostate MSM. The microstate transition matrix for each system was computed at a range of lag times $\tau$ and then diagonalized to recover the corresponding eigenvalues $\lambda_i$ and associated implied time scales $t_i = -\tau / \ln \left| \lambda_i \right|$.~\citep{Wehmeyer2019IntroductionV1.0} The implied time scale plots for the four DNA sequences are presented in SI Figure 2. We observe convergence of the implied time scales very rapidly with lag time $\tau$ for all systems, motivating the construction of high resolution macrostate MSMs at a lag time $\tau$ = 1.2 ns. At this choice of lag time, we recover 5, 4, 2, and 1 implied time scales for the AT-all, GC-end, GC-core, and GC-mid systems, respectively. The identification of $(i-1)$ implied time scales \noteb{(may be 1-2 additional converging, but these correspond to diffusive processes that are irrelevant or unhelpful in determining the hybridization mechanisms. Still, this might come up when looking at the the SI Fig 2} implies the presence of $(i-1)$ slow modes and motivates the coarsening of the system into $i$ macrostates. We estimate these $i$ macrostates by applying PCCA+ spectral clustering to the leading $(i-1)$ eigenvectors of the microstate transition matrix.~\citep{Roblitz2013FuzzyClassification, Weber2018ImplicationsSimulation, Kube2007AConformations} We then estimate the corresponding 6, 5, 3, and 2 macrostate transition matrices $\mathbf{P}$ for the AT-all, GC-end, GC-core, and GC-mid systems, respectively, by projecting the MD simulation trajectories into these discrete macrostates. These macrostate MSMs constitute our terminal kinetic models. We validate the Markovian nature of the four MSMs by subjecting them to the Chapman-Kolmogorov (CK) test \citep{Noe2009ConstructingSimulations,Phys2011MarkovValidation,  Wehmeyer2019IntroductionV1.0}. This test asserts that the transition matrix for a Markovian (i.e., memoryless) MSM constructed at a lag time $\tau$ should satisfy the condition $\mathbf{P}(k \tau) = \mathbf{P}^k(\tau)$, which states that $k$ successive applications of the transition matrix constructed at a lag time $\tau$should be equivalent to a single application of the transition matrix constructed at a lag time $k \tau$. We present in SI Figure 3 the CK tests for each DNA sequence to demonstrate that the $\tau$ = 1.2 ns models performs excellently in predicting transition probabilities out to $k \tau$ = 7.2 ns, validating the Markovian nature and kinetic validity of the four kinetic models.
